The effects of linkage disequilibrium on maintaining balanced polymorphisms in seasonal adaptation

### 2019-11-30 writing/manuscript.Rmd

---
title: "Seasonal Adaptation"
author: \emph{Thomas O'Leary, Csenge Petak, Alison Hall}
date: "11/23/2019"
output: pdf_document
bibliography: refs.bib
header-includes:
   - \usepackage{graphicx}
---

@@

### Basic Model

The method for calculating fitness is based on the work of @Wittmann2017 who originally proposed “segregation lift” as a general mechanism which may maintain balanced polymorphisms in a population under certain conditions.

The model first determines an individual's seasonal score ($z$) according to the equation: write equations later :)

where $n_{homo}$ is the number loci homozygous for the allele particular to that certain season and $d$ is a dominance parameter (which ranges between 0 and 1) and $n_{het}$ is the number of heterozygous loci. A dominance parameter of greater than one ($d > 1$) would correspond to a traditional heterozygote advantage, and therefore is not of interest in this model.

The model we propose differs from the _segregation lift model_ in a few important ways. For one, the loci in our model are located on diploid linked chromosomes, so the homologous pairs can recombine during a simulated meiosis before mating. Expand on why this may be interesting or important… Linked genes not require a $d > 0.5$. Some genes that may be important for seasonal traits (_i.e._ heat shock proteins) are located close to each other on chromosomes and have a high amount of linkage disequilibrium…

@@

| Default Parameters               |      |
|----------------------------------|------|
| Crossover probability            | 0.05 |
| Mutation probability             | 10   |
| Population size                  | 500  |
| Number of loci                   | 100  |
| Exponent of the fitness function | 1    |
| Dominance                        | 0.5  |
| Years                            | 100  |
| Generations per year             | 20   |
| Seasonal balance                 | 2    |

### Simulated seasonal selection

@@

>>>>

---
title: "The effects of linkage disequilibrium on maintaining balanced polymorphisms in seasonal adaptation"
author: \emph{Thomas O'Leary, Csenge Petak, Alison Hall}
output: 
  pdf_document:
    number_sections: true
bibliography: refs.bib
header-includes:
   - \usepackage{graphicx}
   - \usepackage{float}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H',
                      fig.height = 3,
                      fig.width = 5)

require(tidyverse)
source(here::here("functions.R"))
```

@@

## Basic Model

The method for calculating fitness is based on the work of @Wittmann2017 who originally proposed _segregation lift_ as a general mechanism which may maintain balanced polymorphisms in a population under certain conditions.

The model first determines an individual's seasonal score ($z$) according to the equation:

$$
z = n_{homo} + d * n_{het}
$$
where $n_{homo}$ is the number loci homozygous for the allele particular to that certain season and $d$ is a dominance parameter (which ranges between 0 and 1) and $n_{het}$ is the number of heterozygous loci. A dominance parameter of greater than one ($d > 1$) would correspond to a traditional heterozygote advantage, and therefore is not of interest in this model.

The model we propose differs from the _segregation lift_ model in a few important ways. For one, the loci in our model are located on diploid linked chromosomes, so the homologous pairs can recombine during a simulated meiosis before mating. Expand on why this may be interesting or important… Linked genes not require a $d > 0.5$. Some genes that may be important for seasonal traits (_i.e._ heat shock proteins) are located close to each other on chromosomes and have a high amount of linkage disequilibrium…

@@

| Default Parameters               |           |
|----------------------------------|-----------|
| Crossover probability            | 0.05      |
| Mutation probability             | $10^{-4}$ |
| Population size                  | 500       |
| Number of loci                   | 100       |
| Exponent of the fitness function | 1         |
| Dominance                        | 0.5       |
| Years                            | 100       |
| Generations per year             | 20        |
| Seasonal balance                 | 2         |

## Simulated seasonal selection

@@

## Crossover rate

## Epistasis

This is how you reference Figure \ref{epi_density} in the text.

```{r, fig.cap = 'Generations \\label{epi_density}', message = FALSE, echo = FALSE}
setwd(here::here("results/epistasis"))

# load all files and filter only the final generation
all_tbl <- list.files() %>% 
  map_df(~read_plus(.))

tbl_max_gen <- all_tbl %>%
  split(.$filename) %>%
  map(function(.data){filter(.data, genz == max(genz))}) %>%
  do.call("rbind", .)

# plot the density distribution
ggplot(tbl_max_gen, aes(x = freqs, fill = filename)) +
  geom_density(alpha = 0.2) + 
  scale_fill_manual(values = c(rep("#e25dbb", 5), 
                               rep("#6cdf4c", 5),
                               rep("#258bd2", 5)),
                    breaks = unique(all_tbl$filename)[c(1,6,11)],
                    # legend name
                    name = "Epistasis",
                    # values tested
                    labels = c("0.5", "1", "2")) +
  ylab("Density") + 
  xlab("Allele Frequency") +
  theme_classic()

```

## Dominance

## Generations per season

This is how you reference Figure \ref{gen_density} in the text.

```{r, fig.cap = 'Generations \\label{gen_density}', message = FALSE, echo = FALSE}
setwd(here::here("results/generations"))

# load all files and filter only the final generation
all_tbl <- list.files() %>% 
  map_df(~read_plus(.))

tbl_max_gen <- all_tbl %>%
  split(.$filename) %>%
  map(function(.data){filter(.data, genz == max(genz))}) %>%
  do.call("rbind", .)

# plot the density distribution
ggplot(tbl_max_gen, aes(x = freqs, fill = filename)) +
  geom_density(alpha = 0.2) + 
  scale_fill_manual(values = c(rep("#e25dbb", 5), 
                               rep("#6cdf4c", 5),
                               rep("#258bd2", 5)),
                    breaks = unique(all_tbl$filename)[c(1,6,11)],
                    # legend name
                    name = "Generations\nper year",
                    # values tested
                    labels = c("10", "20", "50")) +
  ylab("Density") + 
  xlab("Allele Frequency") +
  theme_classic()

```

## Seasonal balance

## Population size
